---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
#### 1. Load the data
```{r results='hide',warning=FALSE,message=FALSE}
data <- read.csv(unz("activity.zip", "activity.csv"), sep=",",stringsAsFactors=FALSE)
```

#### 2. Transform the data into _data.table_
```{r results='hide',warning=FALSE,message=FALSE}
library(data.table)
dt <- data.table(data);rm(data);gc()
```
```{r echo=FALSE}
dt
```

## The mean total number of steps taken per day.
#### 1. The histogram of the total number of steps taken each day.
```{r}
library(ggplot2)
dt.DailyStep <- dt[,.(DailyStep=sum(steps)),by=date]
qplot(DailyStep, data=dt.DailyStep, geom="histogram",binwidth=1000)
```

#### 2. The mean and median total number of steps taken per day.
```{r}
dt.DailyStep[,.(Mean=mean(DailyStep,na.rm=T),Median=median(DailyStep,na.rm=T))]
```

## The average daily activity pattern.
#### 1. Plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r}
dt.DailyActivity <- dt[,.(AverageSteps=mean(steps,na.rm=T)),by=interval]
qplot(x=interval,y=AverageSteps,data=dt.DailyActivity,geom="line") +
  ylab("Average number of steps")
```

#### 2. The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps:  
```{r}
dt.DailyActivity[order(AverageSteps,decreasing = T)[1],]
```

## Imputing missing values
#### 1. The total number of missing values in the dataset.
```{r}
nrow(dt[which(is.na(steps))])
```


#### 2. Fill in all of the missing values in the dataset using the mean for that 5-minute interval and create a new dataset correspondingly.

```{r, warning=FALSE}
dt.fill <- dt
dt.fill[which(is.na(steps)),]$steps <- dt.DailyActivity[interval %in% dt[which(is.na(steps)),interval],AverageSteps]
dt.fill
```
**Comments: Since the variable 'steps' in the raw data set is of type _Integer_, I filled also integers (by taking round of the means).**

#### 3. The histogram of the total number of steps taken each day, the mean and median total number of steps taken per day. 
```{r}
dt.fill.DailyStep <- dt[,.(DailyStep=sum(steps)),by=date]
qplot(DailyStep, data=dt.fill.DailyStep, geom="histogram",binwidth=1000) +
  xlab("Number of steps taken (daily)") +
  ylab("Frequency")
dt.fill.DailyStep[,.(Mean=mean(DailyStep,na.rm=T),Median=median(DailyStep,na.rm=T))]
```
**According to the graph above, we can see that there are no obvious differences by filling the missing values.**

## Activity patterns at weekdays and weekends.
#### 1. Create a new factor variable in the dataset with two levels: 'weekday' and 'weekend' indicating whether a given date is a weekday or weekend day.
```{r, results='hide'}
dt[,WeekDay:=weekdays(as.Date(dt$date))]
dt[WeekDay=="Sunday" | WeekDay=="Saturday",WD:="weekend"]
dt[!(WeekDay=="Sunday" | WeekDay=="Saturday"),WD:="weekday"]
dt[,WeekDay:=NULL]
```
```{r, echo=FALSE}
dt
```
#### 2. The panel plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days.
```{r}
dt.DailyActivity.WD <- dt[,.(AverageSteps=mean(steps,na.rm=T),WD=WD),by=interval]
qplot(x=interval,y=AverageSteps,data=dt.DailyActivity.WD,geom="line") +
  facet_wrap(~ WD,nrow=2) +
  ylab("Average Number of Steps") +
  theme(strip.background=element_rect(fill="pink"))
```
